---
title: "CHL5233: Quarto in-class assignment"
author: "Ruotong Cheng"
format: pdf
geometry:
  - top=30mm
  - left=20mm
  - right=20mm
  - bottom=30mm
---

For this in-class assignment, you will work in a group to create your **individual** Quarto document. Please consult your classmates and the [Tutorial: Authoring](https://quarto.org/docs/get-started/authoring/rstudio.html) webpage by [Quarto](https://quarto.org/). You can also use the Internet including ChatGPT. The TA and I are also available for any questions.

When you are finished, render the document to PDF. Please carefully **proof read** and make edits as needed. **Commit** and **push** the edits to this file and the new PDF file to your **new repository**.

## Create a **new** repository

- Under your GitHub account, create a **new** repository called `armed_conflict`
  - Throughout the course, we will use this repository to **reproduce** and **replicate** the analysis in the paper
- Create **two** folders
  - `data`
  - `reports`
- Save `countries.txt` and `regions.txt` into `data/raw/` subfolder
- Save `quarto_inclass.qmd` into `reports/` folder

### Example of data science organization

```text
armed_conflict/
├── README.md
├── data/
│   ├── raw/
│   └── processed/
├── scripts/
│   ├── 01_import.R
│   ├── 02_clean.R
│   └── 03_analysis.R
├── outputs/
│   ├── figures/
│   └── tables/
└── reports/
    └── analysis.qmd
```

## Edit this file to create a PDF document using Quarto

The R code chunk below includes the R packages required to knit this document. This is often called a **setup** code. We want the code to run, but we don't want the final document to show the code or results. Edit the code chunk so that the code will run but the code and results (including all warnings and messages) are hidden from the final PDF output.

```{r}
#| include: false
library(tidyverse)
library(knitr)
library(kableExtra)
```

For this exercise, we use two data sets. `countries.txt` contains the list of countries used in the paper *Implications of armed conflict for maternal and child health: A regression analysis of data from 181 countries for 2000-2019* and each country's three-digit ISO code. `regions.txt` contains countries and areas of the world, their ISO country codes, regions, sub-regions, and numeric region codes.

### First

Read in `countries.txt` and create the `countries` data set. Also display the number of rows and columns.

```{r}
countries <- read.table(
  "../data/raw/countries.txt",
  header = TRUE
)

dim(countries)
```

There are `r nrow(countries)` countries included in the `countries` data set.

### Second

Read in `regions.txt` and display the number of rows and columns.

```{r}
regions <- read.table(
  "../data/raw/regions.txt",
  header = TRUE,
  fileEncoding = "latin1"
)

dim(regions)
```

There are `r nrow(regions)` countries and areas included in the `regions` data set.

### Third

Subset the `regions` data set to only include countries from the `countries` data set. The two files are matched using the three-digit ISO country code.

```{r}
regions_subset <- regions %>%
  filter(country_code %in% countries$ISO)

dim(regions_subset)
```

After matching by ISO country code, all `r nrow(regions_subset)` countries are retained.

## Tabulate countries by region and sub-region

Use `group_by()` and `summarise()` to calculate the number of countries in each region. Use `kable()` from the **knitr** package to present the results with column headers `Region` and `Count`.

```{r}
region_table <- regions_subset %>%
  group_by(region) %>%
  summarise(Count = n())

kable(
  region_table,
  col.names = c("Region", "Count")
)
```

Next, calculate the number of countries in each sub-region and create a similar table.

```{r}
subregion_table <- regions_subset %>%
  group_by(sub_region) %>%
  summarise(Count = n()) %>%
  arrange(desc(Count))

kable(
  subregion_table,
  col.names = c("Sub-region", "Count")
)
```

## Data visualization using ggplot

Use `geom_bar()` within `ggplot()` to display the number of countries in each sub-region.

The figure should:

1. Colour-code each sub-region by `region`.
2. Use horizontal bars, with sub-region names on the y-axis.
3. Order the bars so that the sub-region with the most countries appears at the top and the sub-region with the fewest countries appears at the bottom.

```{r}
plot_data <- regions_subset %>%
  mutate(
    sub_region = forcats::fct_rev(
      forcats::fct_infreq(sub_region)
    )
  )

ggplot(
  plot_data,
  aes(x = sub_region, fill = region)
) +
  geom_bar() +
  coord_flip() +
  labs(
    x = "Sub-region",
    y = "Count",
    fill = "Region"
  ) +
  theme_minimal()
```
